An alternative MIP formulation for the Military Flight and Maintenance Planning problem

### 2019-11-22 latex/ROADEF2020/article.tex
changed Nicolas reference affiliation.
changes made by Nicolas. As submitted.

\author{Franco Peschiera\inst{1}, Alain Ha{\"i}t\inst{1}, Olga Batta{\"i}a\inst{2}, Nicolas Dupin\inst{3}}

\institute{
ISAE-SUPAERO, Universit{\'e} de Toulouse, France \\
\email{\{franco.peschiera,alain.hait\}@isae-supaero.fr}\\
KEDGE Business School, France \\
\email{olga.battaia@kedgebs.com} \\
Laboratoire de Recherche en Informatique, France \\
\email{dupin@lri.fr}
}

@@

\keywords{optimization, planning, military, maintenance}

\section{Introduction}
    The Military Flight and Maintenance Planning Problem assigns missions and schedules maintenance operations (checks) for military aircraft. It has been studied in short, medium and long-term planning horizons (\cite{Cho2011,Kozanidis2008,Verhoeff2015,PeschieraR2018}) and it has been proved NP-Hard in its longterm variant by \cite{Peschiera2019}. The present model is an alternative formulation to that of \cite{Peschiera2019} where mission assignments and maintenance cycles are modeled as start-stop assignment.

    The problem consists in assigning aircraft $i \in \mathcal{I}$ to a set $j \in \mathcal{J}$ of missions while scheduling checks over a time horizon divided into $t \in \mathcal{T}$ periods. Missions have known start and end dates; and requires $R_j$ aircraft to fly $H^\prime_{jtt'}$ when assigned between periods $t$ and $t'$. Each aircraft $i$ can be assigned to only missions $j \in \mathcal{JI}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$. Each period $t$ has $j \in \mathcal{JT}_t$ active missions.

    Each aircraft $i$ needs a check after $H^{max}$ flight hours or less; has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon; can be assigned its first check during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, if assigned a first check in period $t$, it can have its second check during periods $t \in \mathcal{T}^{M+}_{t}$. $\mathcal{TTT}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{TTJ}_{jt}$ is the set of mission $j$ patterns that make aircraft be in mission in period $t$; $\mathcal{JTT}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.

@@

@@

    The objective is to delay the second check as much as possible or avoid doing it altogether.
    \begin{align}
        & \text{Max}\; 
        \sum_{
                i \in \mathcal{I}, t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_t
            } m_{itt'} \times t'
        \label{eq:objective1}
    \end{align}

    \begin{align}
        % maximum capacity1:
        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} \leq C^{max} 
                & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Constraints (\ref{eq:capacity1}) limit the number of simultaneous checks. Constraints (\ref{eq:missionres}) enforce aircraft mission requirements. Constraints (\ref{eq:state}) restrict each aircraft to at most one assignment each period.

    \begin{align}
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2T}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2T} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours3}
    \end{align}

    Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of each aircraft before the first check, between checks and after the second check.

    \begin{align}
        & \sum_{t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_{t}} m_{itt'} =  1 
          & i \in \mathcal{I}\label{eq:num_maint}
    \end{align}

    Constraints (\ref{eq:num_maint}) require a check pattern assignment for each aircraft.

@@

    The present work proposes a new model for the Military Flight and Maintenance Problem. The formulation results in better lower bounds (LP-relaxation) compared to \cite{Peschiera2019} and lends itself to pattern prediction forecasting models, which can be used to reduce the size of the model.

@@

\bibliographystyle{plain}
% \selectlanguage{english}
\bibliography{./../biblio/MFMP,./../biblio/FrancoPeschiera}

>>>>

\author{Franco Peschiera\inst{1}, Nicolas Dupin\inst{2}, Olga Batta{\"i}a\inst{3}, Alain Ha{\"i}t\inst{1}}

\institute{
ISAE-SUPAERO, Universit{\'e} de Toulouse, France \\
\email{\{franco.peschiera,alain.hait\}@isae-supaero.fr}\\
KEDGE Business School, France \\
\email{olga.battaia@kedgebs.com} \\
LRI, Universit\'e Paris-Saclay, France \\
\email{dupin@lri.fr}
}

@@

\keywords{optimization, planning, military aircraft, maintenance}

\section{Introduction}
    The Military Flight and Maintenance Planning (MFMP) Problem assigns missions and schedules maintenance operations (checks) for military aircraft. It has been studied in short, medium and long-term planning horizons (\cite{Cho2011,Kozanidis2008,Verhoeff2015,PeschieraR2018}) and it has been proved NP-Hard in its long-term variant by \cite{Peschiera2019}. The present model is an alternative formulation to that of \cite{Peschiera2019} where mission assignments and maintenance cycles are modeled as start-stop assignment.

    The problem consists in assigning aircraft $i \in \mathcal{I}$ to a set $j \in \mathcal{J}$ of missions while scheduling checks over a time horizon divided into $t \in \mathcal{T}$ periods. Missions have known start and end dates and require $R_j$ aircraft to fly $H^\prime_{jtt'}$ hours when assigned to mission $j$ between periods $t$ and $t'$. Each aircraft $i$ can be assigned to only missions $j \in \mathcal{J}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$ if not undergoing a check or a mission. Each period $t$ has $j \in \mathcal{J}_t$ active missions.

    Each aircraft $i$ needs a check after $H^{max}$ flight hours or less and has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon. Only two checks can be assigned in the planing horizon: the first check is assigned during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, a second check can be assigned during periods $t \in \mathcal{T}^{M+}_{t}$, if the first check was done in $t$. $\mathcal{T}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{J}_{jt}$ is the set of mission patterns that make aircraft be in mission $j$ in period $t$; $\mathcal{J}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.

@@

    The following variables are defined in this new Integer Linear Formulation (ILP):

@@

    The objective is to delay the second check as much as possible.

    \begin{small}
         \begin{align}
        & \text{Max}\; 
        \sum_{
                i \in \mathcal{I}, t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_t
            } m_{itt'} \times t'
        \label{eq:objective1}\\
%     \end{align}
%         \begin{align}
        % maximum capacity1:
        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}_{t}} m_{it_1t_2} \leq C^{max} 
                & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{(t_1, t_2) \in \mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
%     \end{align}
% 
% 
%     \begin{align}
        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{it_2T}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2T} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours3}\\
%     \end{align}
% 
%    
%     \begin{align}
        & \sum_{t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_{t}} m_{itt'} =  1 
          & i \in \mathcal{I}\label{eq:num_maint}
    \end{align}
        \end{small}
    Constraints (\ref{eq:capacity1}) limit the number of simultaneous checks,  (\ref{eq:missionres}) enforce aircraft mission requirements and  (\ref{eq:state}) restrict each aircraft to at most one assignment each period.
    Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of each aircraft before the first check, between checks and after the second check. Constraints (\ref{eq:num_maint}) require a check pattern assignment for each aircraft.

@@

    The present work proposes a new model for the MFMP Problem,
    which can be seen as an extended ILP formulation  from \cite{Peschiera2019},  with a     twice-time index formulation.
    This new formulation  induces more variables, but  results in better lower bounds (LP-relaxation) compared to \cite{Peschiera2019}.
   Generic  primal heuristics implemented by MILP solvers  are less efficient with this new formulation, it is explained with the increasing number of variables. To improve both primal and dual bounds of \cite{Peschiera2019}, perspectives are to implement specific matheuristics designed for the extended MILP formulation, using also Machine Learning predictions to guide the search of primal solutions.

%     and lends itself to pattern prediction forecasting models, which can be used to reduce the size of the model.

@@

% \bibliographystyle{plain}
% % \selectlanguage{english}
% %\bibliography{./../biblio/MFMP,./../biblio/FrancoPeschiera}
% \bibliography{MFMP,FrancoPeschiera}
\small
\begin{thebibliography}{1}

\bibitem{Cho2011}
P.~Cho.
\newblock {\em Optimal Scheduling of Fighter Aircraft Maintenance}.
\newblock PhD thesis, MIT, 2011.

\bibitem{Kozanidis2008}
G.~Kozanidis.
\newblock {A Multiobjective Model for Maximizing Fleet Availability under the
  Presence of Flight and Maintenance Requirements}.
\newblock {\em Journal of Advanced Transportation}, 43(2):155--182, 2008.

\bibitem{Peschiera2019}
F. Peschiera, A. Ha{\"i}t, N. Dupin, and O. Batta{\"i}a.
\newblock Long term planning of
military aircraft flight and maintenance operations. 
\newblock Technical report, ISAE-SUPAERO,
Universit\'e de Toulouse, 2019.

\bibitem{PeschieraR2018}
F. Peschiera, A. Ha{\"i}t, N. Dupin, and O. Batta{\"i}a.
\newblock Maintenance planning on french military aircraft operations.
\newblock In {\em Congr{\`e}s annuel de la ROADEF}, 2018.

\bibitem{Verhoeff2015}
M.~Verhoeff, W.~J~C Verhagen, and R.~Curran.
\newblock {Maximizing operational readiness in military aviation by optimizing
  flight and maintenance planning}.
\newblock {\em Transportation Research Procedia}, 941--950, 2015.

\end{thebibliography}
